---
title: "Final Project"
author: "Ryuhei Shida"
date: 2020-03-20
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, warning=FALSE, message=FALSE, include=FALSE}
library(tidyverse)
library(rvest)
library(dplyr)
library(lubridate)
```
# Scrape Data

I scraped European football data from an online database using the rvest package.

Using rvest, I got data about every team in the "top 4 leagues" of European Football, which are England, Germany, Italy, and Spain.

Source of Material is http://www.footballsquads.com. Material: © FootballSquads.com, 1999 - 2020, All Rights Reserved

Scraping was a complex process, where I first needed to get the links to the indivual teams for each league.

```{r, eval=FALSE, include=FALSE}
# Stores html source of every league
leagues_html = list()

leagues_html[[1]] <- read_html("http://www.footballsquads.co.uk/eng/2019-2020/engprem.htm")
leagues_html[[2]] <- read_html("http://www.footballsquads.co.uk/ger/2019-2020/gerbun.htm")
leagues_html[[3]] <- read_html("http://www.footballsquads.co.uk/italy/2019-2020/seriea.htm")
leagues_html[[4]] <- read_html("http://www.footballsquads.co.uk/spain/2019-2020/spalali.htm")
```


```{r, eval=FALSE, include=FALSE}
# Stores links to teams per league
leagues_teams = list()

for(i in 1:length(leagues_html)) {
  leagues_teams[[i]] <- leagues_html[[i]] %>% 
    html_nodes("h5") %>% 
    html_nodes("a") %>% 
    html_attr("href")
}
```

Then I went through each league's teams, and scraped the table on the site.

I had to add the team name as a column, and remove players who were no longer at their club.

```{r, eval=FALSE, include=FALSE}
base_urls = list()
base_urls[1] = "http://www.footballsquads.co.uk/eng/2019-2020/"
base_urls[2] = "http://www.footballsquads.co.uk/ger/2019-2020/"
base_urls[3] = "http://www.footballsquads.co.uk/italy/2019-2020/"
base_urls[4] = "http://www.footballsquads.co.uk/spain/2019-2020/"
```


```{r, warning=FALSE, eval=FALSE}

# Stores players from every league
leagues_allplayers = list()

# Go through each league
for(league in 1:length(leagues_teams)) {

  team_names = list()
  team_squads = list()
  
  for(team in 1:length(leagues_teams[[league]])) {
    s = html_session(base_urls[[league]]) %>%
      jump_to(leagues_teams[[league]][team]) %>%
      read_html()
      
    team_names[[team]] <- s %>% 
      html_nodes("h2") %>% 
      html_text
    
    
    team_squads[team] <- s %>% 
      html_table(header = TRUE)
  
  }
  
  # For each team in current league
  for(i in 1:length(team_names)) {
    # Add team name
    team_squads[i] <- lapply(team_squads[i],
                             function(x) cbind(x, Team = team_names[[i]]))
    
    # Remove "Players no longer at this club"
    team_squads[i] <- lapply(team_squads[i],
                             head,
                             n = which(as.data.frame(team_squads[i])$Number == "Players no longer at this club")-1)
  
    leagues_allplayers[[league]] <- bind_rows(team_squads)

  }
}

```

After I had data frames for each league, I wrote a function to add the league name as a column and remove numbers that are missing.

```{r, eval=FALSE, include=FALSE}
league_names = c("England", "Germany", "Italy", "Spain")
```

```{r, eval=FALSE}
# Function to 
assign_and_remove <- function(index) {
  leagues_allplayers[[index]]["League"] <- league_names[index]
  leagues_allplayers[[index]] <- leagues_allplayers[[index]][!(is.na(leagues_allplayers[[index]]$Name) | leagues_allplayers[[index]]$Name==""), ]
  leagues_allplayers[[index]]
}
```

```{r, eval=FALSE, include=FALSE}
assign_and_remove(1)
```

```{r, eval=FALSE, include=FALSE}
assign_and_remove(2)
```

```{r, eval=FALSE, include=FALSE}
assign_and_remove(3)
```

```{r, eval=FALSE, include=FALSE}
assign_and_remove(4)
```

# Map Players to Birth Place

Using the `tidygeocoder` package, we geocode players from each league.

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(sf)
library(mapview)
library(tidygeocoder)
```

```{r, eval=FALSE}
# Geocode
get_coords <- function(index) {
  leagues_allplayers[[index]] %>% 
    geocode(`Birth Place`, method = "osm")
}
```

```{r, eval=FALSE, echo=FALSE}
geocoded = list()
```



```{r, eval=FALSE, include=FALSE}
# Write geocoded data frames to csv files
write.csv(file="geocoded1.csv", x=geocoded[[1]], row.names = FALSE)
write.csv(file="geocoded2.csv", x=geocoded[[2]], row.names = FALSE)
write.csv(file="geocoded3.csv", x=geocoded[[3]], row.names = FALSE)
write.csv(file="geocoded4.csv", x=geocoded[[4]], row.names = FALSE)
```


```{r, echo=FALSE}
geocoded = list()
# Read data from csv files to save time

geocoded[[1]] <- read.table("geocoded1.csv", header = TRUE, sep = ',', encoding = 'latin1')
geocoded[[2]] <- read.table("geocoded2.csv", header = TRUE, sep = ',', encoding = 'latin1')
geocoded[[3]] <- read.table("geocoded3.csv", header = TRUE, sep = ',', encoding = 'latin1')
geocoded[[4]] <- read.table("geocoded4.csv", header = TRUE, sep = ',', encoding = 'latin1')
```

The geocoded list holds the data frames of players in each of the 4 leagues.


I finally put the data for all 4 leagues together, now that they are mostly prepared for analysis.

```{r, warning=FALSE}
top4leagues <- bind_rows(list(geocoded[[1]],
                              geocoded[[2]],
                              geocoded[[3]],
                              geocoded[[4]]))
```


```{r}
# Remove numbers that are missing
top4leagues <- top4leagues[!(is.na(top4leagues$Name) | top4leagues$Name==""), ]
```

```{r}
head(top4leagues)
```

```{r, eval=FALSE, echo=FALSE}
# Save to csv file
write.csv(file="top4leagues.csv", x=top4leagues, row.names = FALSE)
```



I manually added information for some players with missing coordinate data.

```{r, eval=FALSE}
# Add information for some notable missing players
top4leagues[top4leagues$Name == "Mohamed Salah", "lat"] = 30.940491
top4leagues[top4leagues$Name == "Mohamed Salah", "long"] = 30.830231
top4leagues[top4leagues$Name == "Bukayo Saka", "lat"] = 51.525
top4leagues[top4leagues$Name == "Bukayo Saka", "long"] = -0.3325
```



```{r}
# Keep players with lat and long not na
top4leagues2 <- top4leagues[complete.cases(top4leagues[12:13]), ]
```


```{r, echo=FALSE}
england <- top4leagues2 %>% 
  filter(League == "England")

germany <- top4leagues2 %>% 
  filter(League == "Germany")

italy <- top4leagues2 %>% 
  filter(League == "Italy")

spain <- top4leagues2 %>% 
  filter(League == "Spain")
```

I created data frames for each country again, and plotted where the players who played in each respective league were born. The color of the dots show their position.

### England

```{r}
locations_sf <- st_as_sf(england, coords = c("long", "lat"), crs = 4326)
mapview(locations_sf, zcol=c("Pos"))
```

### Germany

```{r}
locations_sf <- st_as_sf(germany, coords = c("long", "lat"), crs = 4326)
mapview(locations_sf, zcol=c("Pos"))
```

### Italy

```{r}
locations_sf <- st_as_sf(italy, coords = c("long", "lat"), crs = 4326)
mapview(locations_sf, zcol=c("Pos"))
```

### Spain

```{r}
locations_sf <- st_as_sf(spain, coords = c("long", "lat"), crs = 4326)
mapview(locations_sf, zcol=c("Pos"))
```

- Players from South Africa tend to play in England
- Asian Players tend to be forwards in England, midfielders in Germany and Spain, and defenders in Italy.

# Birthdays

```{r, include=FALSE}
top4leagues3 <- top4leagues2
```

```{r}
top4leagues3$bd <- dmy(top4leagues3$Date.of.Birth)
top4leagues3$bd_year <- year(top4leagues3$bd)
```


```{r}
top4leagues3 %>% 
  ggplot(aes(x = bd_year)) +
  geom_density(stat = "count") +
  ggtitle("Birthday of Active Players")
```


```{r}
summary(top4leagues3$bd)
```

# Count per team

```{r}
top4leagues2 %>% 
  group_by(League, Team) %>% 
  summarise(`Player Count` = length(Team)) %>% 
  arrange(desc(`Player Count`)) %>% 
  head(10)
```

Liverpool, Tottenham Hotspur, and Manchester United have the highest player count

# Height and Weight per position

```{r, warning=FALSE}
top4leagues2 %>% 
  ggplot(aes(x = Height, color = Pos)) +
  geom_density() +
  ggtitle("Height Distribution by Position")
```


```{r}
tapply(top4leagues2$Height, top4leagues2$Pos, summary)
```

Goalkeepers have the highest height.

```{r, warning=FALSE}
top4leagues2 %>% 
  ggplot(aes(x = Height, color = Pos)) +
  geom_density() +
  ggtitle("Weight Distribution by Position")
```

```{r}
tapply(top4leagues2$Weight, top4leagues2$Pos, summary)
```

Goalkeepers have a much larger weight than other positions.